---
title: "Goose Data, Inter-annual variation maps"
---

```{r, include=FALSE}
#'load knitr
library(knitr)

#'suppress all code output but run code
opts_chunk$set(echo = FALSE,
               root.dir = normalizePath("~/git/thesis/"))
opts_knit$set(root.dir = normalizePath("~/git/thesis/"))
```

```{r load_shp, message=FALSE, include=FALSE}
#'load some polygons
library(sp);library(maptools); library(rgdal); library(raster)
nl = readOGR("vector/ne_10m_admin_0_boundary_lines_land.shp")
coast = readOGR("vector/ne_10m_land.shp")
lakes = readOGR("vector/ne_10m_lakes.shp")
```

```{r load_clean_data, include=FALSE}
#'load data
geese = read.csv("data.goose.clean.csv", row.names = 1)
```


```{r summarise_region_year}
#'use ddply to get mean flocksize, percjuv and famsize by zone and by region
#'load plyr
library(plyr)

zone_data = ddply(geese, c("zone","Breeding_year"), summarise, juvprop = mean(propjuv, na.rm=T),
      flocksize = mean(flocksize, na.rm=T), 
      famsize = mean(meanfam))
```

```{r get_province_codes}
#import place codes
provinces = read.csv("province_coords.csv")[,c(3:5)]
```

```{r get_zone_codes}
#'choose zone codes from province codes
zones = provinces[provinces$V2 %in% c("DU","LER","FR","ZL"),]
#'change region names to zones
zones$V2 = c("Rhinelands","IJsselmeer","East Frisia","Southwest")
#'merge with zone data
zone_data = merge(zone_data, zones, by.x = "zone",by.y = "V2")

#make zonedata spatial
coordinates(zone_data) = ~lon+lat
```

# Juvenile proportion in zones over years

```{r map_juvprop_split_by_year, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, fig.height=10, fig.width=15}
#'load tmap
library(tmap)
tm_shape(coast,bbox = extent(zone_data)+0.5)+
  tm_polygons(col = "grey90")+
tm_shape(nl)+
  tm_lines()+
tm_shape(zone_data)+
  tm_symbols(scale = 6,border.lwd = 0.1, size = "flocksize",
             col = "juvprop", legend.size.show= F,
            legend.col.show = T, style = "cont",
            #breaks = seq(0,70,10),
            title.size = "Mean flock size", 
            title.col = "Mean juvenile %")+
 tm_text("zone", size = 0.3, just = c("right","bottom"))+
  
tm_facets(by = "Breeding_year", free.scales=FALSE, ncol = 4)
```

# Family sizes at geocoded sites over years

```{r import_fam_data, include=FALSE}
#'import family size
family_sites = read.csv("family_sites.csv")

#'make spatial
coordinates(family_sites) = ~lon+lat
```

```{r map_famsize_locs_years, echo=FALSE, fig.fullwidth=TRUE, fig.height=10, fig.width=15, message=FALSE, warning=FALSE}
#'tmapping facet by year
tm_shape(coast,bbox = extent(zone_data)+0.5)+
  tm_polygons(col = "grey80")+
tm_shape(nl)+
  tm_lines()+
tm_shape(family_sites)+
  tm_symbols(scale = 6, col = "meanfam",
             size = "flocksize", jitter = 1,
             style = "cont", legend.size.show = F)+
  tm_facets(by = "Breeding_year", ncol = 4)
```

From the spatial pattern seen in the data, I classify the goose data into four major regions: The Rhinelands, encompassing the region around Duisburg, Wesel, Cleves and Nijmegen, the North East, or the regions marked Leer, Groningen and Drenthe, the Southwest, with Zeeland and South Holland, and the IJsselmeer, comprising the regions marked Frisia, Flevoland and North Holland.

```{r classify_region}
#'work with family_records.csv

family_records = read.csv("family_records.csv")

#'assign new zones
family_records$zone = ifelse(family_records$Region %in% c("DU","KLE","RG","WES","WG"),
       "Rhinelands",
       ifelse(family_records$Region %in% c("LER","DR","GR"),
              "East Frisia",
       ifelse(family_records$Region %in% c("FR","FL","NH"),
              "IJsselmeer", 
              ifelse(family_records$Region %in% c("ZL","ZH"),
                 "Southwest", "Other"))))

#'table of frequencies
freq_zone = count(family_records, c("zone", "Breeding_year"))

#merge zone coords and make spatial
family_records = merge(family_records, zones, by.x = "zone",
                     by.y = "V2")
coordinates(family_records) = ~lon+lat
```

```{r map_famsize_zones_years, echo=FALSE, fig.fullwidth=TRUE, fig.height=10, fig.width=15, message=FALSE, warning=FALSE}
#'zone as factor
family_records$zone = as.factor(family_records$zone)

#'tmapping facet by year

tm_shape(coast,bbox = extent(zone_data)+1)+
  tm_polygons(col = "grey80")+
tm_shape(nl)+
  tm_lines()+
tm_shape(family_records)+
  tm_symbols(scale = 4.5, col = "meanfam",
             jitter = 2, size = "flocksize",
             style = "cont", legend.size.show = T,
             shape = "zone")+
  tm_facets(by = "Breeding_year", ncol = 4)
```