---
title: Cleaning Data
output: 
  tufte::tufte_handout
---

```{r, include=FALSE}
#'load knitr
library(knitr)

#'suppress all code output but run code
opts_chunk$set(eval = FALSE)
opts_knit$set(root.dir = "~/git/thesis/")
```

# Import data from excel file, set column types

```{r load data, include=FALSE}
#'load libs
library(readxl)
library(RColorBrewer)

#'import data
data.raw.goose = read_excel("~/git/thesis/Age-ratiodata-GWfG-toPratik.xlsx", 
    sheet = "Plain_table", col_types = c("text", 
        "numeric", "numeric", "numeric", 
        "numeric", "text", "text", "text", 
        "text", "numeric", "text", "text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "text", "text", 
        "text"))
```

# Define general spatio-temporal extent

```{r reduce_data}
#'all records in the netherlands or nrw after than 2000

#'first keep only nl or nrw after 2000
goose.clean = data.raw.goose[data.raw.goose$Country %in% c("NL","NRW"),]

goose.clean = goose.clean[goose.clean$Breeding_year > 2000,]
```

# Get time as a new column, define explicit temporal extents

```{r assign_time}
#'fix missing days, if days are missing, assign the 15th of the month
goose.clean$Day = ifelse(is.na(goose.clean$Day),
                        15, goose.clean$Day)

#'collect the contents of the year, month and day columns as a single column of the posixct class. this may need to be modified to date class later since it also assigns a time, which causes problems when the timezone shifts due to DST
goose.clean$time = as.POSIXct(paste(goose.clean$Year,                                    goose.clean$Month,goose.clean$Day),
                             format="%Y %m %d")

#'remove records after may 2017
goose.clean = goose.clean[goose.clean$time < "2017-05-01",]
```

# Fix missing counts

```{r fix_flocksize}
#'fix flock size:
#'if flock size in the field total_flock is not recorded, substitute the field n_sampled. this is based on the assumption that all goose.clean were whitefronts
#'if n_sampled is NA, assign the sum of juveniles and adults as the flock size. same assumption as above holds.
#'finally, if flock size is recorded, leave it as it is.
goose.clean$Total_flock = ifelse(is.na(goose.clean$Total_flock),
        ifelse(is.na(goose.clean$N_sampled),
               goose.clean$N_Adult+goose.clean$N_Juvenile,
               goose.clean$N_sampled),
                        goose.clean$Total_flock)

#'fix juvenile percentage: if the field perc-jv is NA, calculate the juvenile percentage as number of juveniles*100 divided by the field total_flock
#'this is changed now. the juvenile percentage is calculated on whitefronts ONLY!!!
goose.clean$`Perc-JV` = ifelse(is.na(goose.clean$`Perc-JV`),
          (goose.clean$N_Juvenile/goose.clean$N_sampled)*100,
                        goose.clean$`Perc-JV`)

#'check for NAs
apply(goose.clean, 2, function(x){sum(is.na(x))})
```

# Export basic clean data

```{r export_basic_clean_data}
#export to csv

write.csv(goose.clean, "~/git/thesis/code/data.goose.clean.csv")
```

# Add zone as a variable, rename some variables

```{r add_vars}
#'duplicate goose.clean

#'assign zones: if in DU, KLE, RG, WES, WG, assign to Rhinelands, if itn LER, DR, GR, assign to East Frisia, if in FR, FL, NH, assign to IJsselmeer, and if in ZL, ZH assign to Southwest, assing all else to other.
goose.clean$zone = ifelse(goose.clean$Region %in% c("DU","KLE","RG","WES","WG","LI"),
       "Rhinelands",
       ifelse(goose.clean$Region %in% c("LER","DR","GR"),
              "East Frisia",
       ifelse(goose.clean$Region %in% c("FR","FL","NH"),
              "IJsselmeer",
              ifelse(goose.clean$Region %in% c("ZL","ZH"),
                     "Southwest", "Other"))))

#'rename columns to make typing easier
library(plyr)
goose.clean = rename(goose.clean, replace = c("Perc-JV" = "propjuv",
                                  "N_Juvenile" = "juvs",
                                  "Total_flock" = "flocksize",
                                  "Mean_Fam" = "meanfam",
                                  "N_Adult" = "adults"))

#'keep a reduced number of columns
goose.clean = goose.clean[,c("Breeding_year","Region","Site_name",
                 "flocksize", "adults", "juvs", "N_sampled",
                 "propjuv", "Fam1", "Fam2", "Fam3",
                 "Fam4", "Fam5", "Fam6", "Fam7",
                 "Fam8", "Fam9", "Fam10", "meanfam",
                 "time", "zone", "Observer","Food_type")]

#'sum and count the number of families
goose.clean$fams = apply(goose.clean[,c(9:18)], 1, function(x){sum(x, na.rm = T)})

#'subset for family size data
fams = goose.clean[goose.clean$fams > 0,]
```

# Geocode the data

```{r geocoding_and_spatial_sites}
#'merge the dataframe fams which has detailed family count data with geocoded sites, and with geocoded zones
#'get coords
sites.coords = read.csv("coords.csv", row.names = 1)

#'merge the same for families, creating a new object called fams.coords
fams.coords = merge(fams, sites.coords, by.x = "Site_name", by.y = "places")

#'remove sites where no coordinates were found
fams.coords = fams.coords[complete.cases(fams.coords[,c("lon","lat")]),]

#'set spatial extent for both
fams.coords = fams.coords[fams.coords$lat > 0 & fams.coords$lat <56 &
                          fams.coords$lon >0 & fams.coords$lon <10,]
```

```{r coordinates_zones}
#'get the province coordinates
#'preceding code was:
zones = read.csv("province_coords.csv", row.names = 1)
zones = zones[zones$regions %in% c("Leer","Friesland","Duisburg","Rotterdam"),]
zones$zone = c("Rhinelands","IJsselmeer","East Frisia","Southwest")

#'merge goose.clean with the zones
goose.clean = merge(goose.clean, zones[,c("lon","lat","zone")], by.x = "zone", by.y = "zone", all.x = T)
```

# Export data

```{r export_data}
#'write to csv as data.goose.clean
write.csv(goose.clean, "data.goose.clean.csv")

#'write the full family data to csv
write.csv(fams, "fams.csv")

#'write geocoded family data to csv
write.csv(fams.coords, "data.fams.coords.csv")
```

# Expand family data

```{r expand_family_data}
#set all fams from NA to 0
fams[,c(9:18,22)][is.na(fams[,c(9:18,22)])] = 0

#'get expanded fams data, zone, time, etc
fams.expand <- fams[rep(row.names(fams), fams$fams), c(1:8,19:dim(fams)[2])]
#'get the families as a list
famrep = apply(fams[,c(9:18)], 1, function(x){rep(c(1:10), x)})
#'unlist it for a fam sequence
famrep = unlist(famrep)

#'bind with fams.expand
fams.expand$famsize = as.numeric(famrep)
```

# Export non-spatial family data

```{r write_expanded_family_data}
#'write the expanded family data to csv. this data is not geocoded
write.csv(fams.expand, "fams.expand.csv", row.names = F)
```

# Make data spatial

```{r make_expanded_data_spatial}
#'make expanded data spatial
fams.expand.coords = merge(fams.expand, sites.coords, by.x = "Site_name",by.y="places")

#'clean data
fams.expand.coords = fams.expand.coords[fams.expand.coords$lat >0 &
                                   fams.expand.coords$lon > 0 &
                                   fams.expand.coords$lon <=10,]
#'remove NA coords
fams.expand.coords = fams.expand.coords[complete.cases(fams.expand.coords[,c("lon","lat")]),]

#'write to csv
write.csv(fams.expand.coords, "fams.expand.coords.csv")
```

# Export spatial data as shapefiles

```{r export_shapefile}
#'save as shapefiles
library(rgdal)
#'make spatial
coordinates(fams.coords) = ~lon+lat
coordinates(fams.expand.coords) = ~ lon+lat

#'write flock with family counts data to shapefile
writeOGR(fams.coords, "data_shapefiles", layer = "fams.coords",driver = "ESRI Shapefile", overwrite_layer = T)

#'write family expanded data to shapefile
writeOGR(fams.expand.coords, "data_shapefiles",layer = "fams.expand.coords",
         driver = "ESRI Shapefile", overwrite_layer = T)
```

