---
output:
 tufte::tufte_handout
---

# Study area and datasets

```{r, include=FALSE, message=FALSE, warning=FALSE}
#'load knitr
library(knitr)
library(tmap);library(gridExtra);library(ggplot2)

#'suppress all code output but run code
opts_chunk$set(echo = FALSE)
opts_knit$set(root.dir = "~/git/thesis/code")
tufte::tufte_handout()

#'load data
load("goose_datamap.RData")
load("goose_dataplots.Rdata")
```

```{r map1, fig.fullwidth=TRUE, fig.height=6, fig.width=8, warning=FALSE}

map01

```

**Data collection sites in the study area.**
Three datasets used in the study are represented on a map of the main wintering grounds of the North Sea population of Whitefronts. Lines represent coasts (light) and major rivers (dark).

Crosses mark sites (*n* = 64) where the sizes of flocks of Whitefronts, and the numbers of families with at least one juvenile within them were recorded between autumn 2000 and spring 2017.
Triangles mark positions (*n* = 19) from 13 GPS tracked families of geese (3 in 2013, 4 in 2014, 6 in 2016) where individuals left the family (see details in text). Sites where geese with numbered neckbands were observed, and their family sizes counted, between 2000 and 2017, are bounded by a kernel shaded grey ($n_{obs}$ = 10,635, $n_{sites}$ = 8,416).

# Trends in family size

```{r plot_grid1, fig.fullwidth=TRUE, fig.height=6, fig.width=8, warning=FALSE, message=FALSE}

pdf(file = "~/git/thesis/texts/fsize_trends.pdf")
grid.arrange(p1, p2, p3, p4)

dev.off()
```

**Model output for generalised models run on two datasets with family size as response.**

Top row: Model fit and data for family size with longitudinal position, days since first autumn arrivals, summer predation, and flock size as fixed effects. Habitat type, observer, and breeding year are *iid.* random effects. Data used were complete cases of counts (*n* = 34,174) of successful families in flocks. Lines show partial fits for longitude (a), and for days since goose arrivals in autumn (b). Model $\Omega^{2}_0$ = 0.563.

Bottom row: Model fit and data for family size with longitudinal position, days since first autumn arrivals, and summer predation as fixed effects. Individual identity and individual identity nested within breeding year are random effects. Data used were all families with marked geese (*n* = 10,426), and a subset of only successful families (*n* = 3,102). Lines show partial fits for longitude (c), with all families (solid line), and only successful families (dashed line), and for days since arrivals (d). Model $\Omega^{2}_0$ = 0.786, all families; $\Omega^{2}_0$ = 0.661, successful families.

\newpage
# Trends in flock size
```{r plot_grid2, fig.fullwidth=TRUE, fig.height=3, fig.width=8, warning=FALSE}
pdf(file = "~/git/thesis/texts/flock_trends.pdf")
grid.arrange(p5, p6, p7, p8, ncol = 2)
dev.off()
```

**Model output of generalised models run on flock count data.**

Top row: Model fit and data for flock size with longitudinal position, days since first autumn arrivals, and summer predation as fixed effects. Data used were complete cases of flock counts (*n* = 5,700). Lines show partial fits for **(a)** longitude, and **(b)** for days since goose arrivals in autumn . Model $\Omega^{2}_0$ = 0.99.
Bottom row: Model fit line and data for (c) number of families in flocks, with summer predation and longitude as fixed parametric effects and flock size as a smooth term, and breeding year, observer and habitat type as random effects. Data used were family frequency data from flock counts (*n*  = 837). Model $\Omega^{2}_0$ = 0.99. Model fit and data for (d) family size with longitude, flock size, days since first autumn arrivals, and summer predation as fixed effects. Data were family sizes from flock counts (*n* = 34,174). Line shows partial fit for flock size. Model $\Omega^{2}_0$ = 0.562.
Habitat type, observer, and breeding year are *iid.* random effects in both cases.

\newpage
# Trends in juvenile proportion

```{r plot4, fig.fullwidth=TRUE, fig.height=3, fig.width=8, warning=FALSE}
pdf(file = "~/git/thesis/texts/juv_trends.pdf")
grid.arrange(p9, p10, nrow = 1)
dev.off()
```

**Trends in the proportion of first winter birds in flocks of Greater White-fronted geese.**
Model fit and data for juvenile proportion with longitudinal position and summer predation as fixed effects, and days since first autumn arrivals as a smooth term. Data used were complete cases of flock age ratios (*n* = 5,659). Lines show partial fits for **(a)** days since arrivals, and **(b)** flock size. Model $R^2$ = 0.104.

\newpage
# The effect of summer predation on flocks

```{r echo =FALSE, message=FALSE,warning=FALSE, fig.height=6, fig.width=8, fig.fullwidth = TRUE}
pdf(file = "~/git/thesis/texts/pred_trends.pdf", height = 6, width = 8)
grid.arrange(p11, p12, p13, p14, nrow = 2)
dev.off()
```

**Trends in flock size and proportion of first winter juveniles in relation to summer predation.**

Model fits and data for **(a)** flock size with longitude, days since first autumn arrivals, and summer predation as fixed effects. Data used were complete cases of flock counts (*n* = 5,700). Model $\Omega^{2}_0$ = 0.99.
**(b)** Juvenile proportion with flock size, longitude, and summer predation as fixed parametric effects, and days since arrivals as a smooth term. Data used were complete cases of flock age ratios (*n* = 5,659). Model $R^2$ = 0.104.
**(c)** Number of successful families in flocks with longitude, days since autumn arrivals, flock size and summer predation as fixed effects. Data used were family frequencies from flock counts (*n* = 5,659). Model $\Omega^{2}_0$ = 0.99.
**(d)** Family size with flock size, longitude, days since arrivals and summer predation as fixed effects. Data used were family sizes from flock counts (*n* = 34,174), and observations of individual geese (*n* = 10,426). Model $\Omega^{2}_0$ = (*1*) 0.661, successful families in flocks, (*2*) 0.563, successful pairs observed individually, and (*3*) 0.786, all pairs including unsuccessful ones observed individually. Random effects in **(a)**, **(b)** and **(c)** & **(d)**(*2*) are *iid.* breeding year, observer, and habitat type, and in **(d)**(_1, 3_) are goose identity, and goose identity nested within breeding year.
Lines in **(a)**, **(b)** and **(c)** show partial fit for summer predation index. Lines in **(d)** show partial fit of family sizes of (*1*) successful pairs from individual observations, (*2*) successful families from flock counts, and (*3*) all pairs including unsuccessful ones from individual observations. 95% confidence intervals are shaded grey.

