---
title: thesis model plots
---

```{r set_env, message=FALSE, echo=FALSE, warning=FALSE}
#'load knitr
library(knitr);library(plyr);library(dplyr);library(data.table);library(purrr);library(ggplot2)

#'load model libs
library(gam);library(mgcv);library(gamm4);library(itsadug)

#'load variogram libraries
library(sp);library(gstat);library(visreg)

#'suppress all code output but run code
opts_chunk$set(eval = FALSE)
opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE, echo=FALSE, warning=FALSE, eval=TRUE}
source("dataprep.R")
load("goosemods_pres.RData")
load("distmods.Rdata")
load("fsize_coords.rdata")

#'define ggplot axis and grid options
a = theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
b = theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

# Plots: Distance to Kolguyev ~ family size

```{r distkol_fams1, message=FALSE, echo=FALSE, warning=FALSE}
#'get model response for lon
vis.distmod1 = visreg(distmod, "famsize", scale = "response", plot = F)

out.distmod1 = vis.distmod1$fit
```

```{r export_distkol_fams1, eval = FALSE}
#postscript(file = "plot01distfam1.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

#'plot
p1 = ggplot()+
  geom_ribbon(data = out.distmod1, aes(x = famsize, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = out.distmod1, aes(x = famsize, y = visregFit), col = 2)+
  labs(list(x = NULL, y = "Dist. Kolguyev (km)", title = "a"))+
  geom_jitter(data = fams.expand.sub %>% filter(t_since_in < 60), aes(x = famsize, y = distkol), size = 0.1, pch = 1)+

#  geom_segment(data = splits %>% filter(event %in% c("Split")), aes(x = lon, y = fsize+1, xend = lon, yend = 0), lty = 1, lwd = 0.2, col = "grey30")+
 # geom_text(data = splits %>% filter(fam %in% c("Ev","Wo","Ha","Na","Ti","Ma","Ro","An","Jo"), event == "Split"), aes(x=lon, y = fsize+1, label = substr(fam,1,1), angle = 0), hjust = 1, vjust = 0, size = 3, angle = -30)+
 geom_text(aes(x =8, y = 3e3, label = "{chi}^2==1.135"), parse = T, hjust = "inwards")+xlim(0,10)+
  theme_bw()+a
#dev.off()
```

```{r distkol_fam1, message=FALSE, echo=FALSE, warning=FALSE}
#'get model response for lon
vis.distmod1l = visreg(distmod_late, "famsize", scale = "response", plot = F)

out.distmod1l = vis.distmod1l$fit
```

```{r export_distkol_fams1l, eval = FALSE}
#postscript(file = "plot01distfam1l.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

#'plot
p2 = ggplot()+
  geom_ribbon(data = out.distmod1l, aes(x = famsize, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = out.distmod1l, aes(x = famsize, y = visregFit), col = 2)+
  labs(list(x = NULL, y = NULL, title = "b"))+
  geom_jitter(data = fams.expand.sub %>% filter(t_since_in > 60), aes(x = famsize, y = distkol), size = 0.1)+

 geom_text(aes(x =8, y = 3e3, label = "{chi}^2==4.194"), parse = T, hjust = "inwards")+xlim(0,10)+
  theme_bw()+b
#dev.off()
```

```{r distkol_fam2, message=FALSE, echo=FALSE, warning=FALSE}
#'get model response for lon
vis.distmod2 = visreg(distmod2, "famsize", scale = "response", plot = F)

out.distmod2 = vis.distmod2$fit
```

```{r export_distkol_fams2, eval = FALSE}
#postscript(file = "plot01distfam1l.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

#'plot
p3 = ggplot()+
  geom_ribbon(data = out.distmod2, aes(x = famsize, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = out.distmod2, aes(x = famsize, y = visregFit), col = 2)+
  labs(list(x = "Family size", y = "Dist. Kolguyev (km)", title = "c"))+
  geom_jitter(data = geeseorg %>% filter(t_since_in < 60), aes(x = famsize, y = distkol), size = 0.1)+

 geom_text(aes(x =8, y = 3.1e3, label = "{chi}^2==2.007"), parse = T, hjust = "inwards")+xlim(0,10)+
  theme_bw()+a
#dev.off()
```

```{r distkol_fam2l, message=FALSE, echo=FALSE, warning=FALSE}
#'get model response for lon
vis.distmod2l = visreg(distmod2_late, "famsize", scale = "response", plot = F)

out.distmod2l = vis.distmod2l$fit
```

```{r export_distkol_fams2, eval = FALSE}
#postscript(file = "plot01distfam1l.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

#'plot
p4 = ggplot()+
  geom_ribbon(data = out.distmod2l, aes(x = famsize, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = out.distmod2l, aes(x = famsize, y = visregFit), col = 2)+
  labs(list(x = "Family size", y = NULL, title = "d"))+
  geom_jitter(data = geeseorg %>% filter(t_since_in > 60), aes(x = famsize, y = distkol), size = 0.1)+

 geom_text(aes(x =8, y = 3.1e3, label = "{chi}^2==0.27"), parse = T, hjust = "inwards")+ xlim(0,10)+
  theme_bw()+b
#dev.off()
```

# Family size ~ time: SOVON data

```{r modresponse_fam_time}
#'get model response for t_since_in
vis.mod03.time = visreg(mod03, "t_since_in", gg = TRUE, scale = "response", line.par = list(col = 1), plot = F)

#'make data frame
vis.mod03.time.data = vis.mod03.time$fit

#'add t_since_in to splits data
splits$time = as.Date(splits$time)

#'merge mig.red and splits
splits = merge(splits, mig.red, by.x = "time", by.y = "Date")
```

```{r export_fam_time, warning=FALSE, message=FALSE}
#postscript(file = "plot02sovonfamtime.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

p5 = ggplot()+
  geom_ribbon(data = vis.mod03.time.data, aes(x = t_since_in, ymin = visregLwr, ymax = visregUpr), fill = "grey90")+
  geom_line(data = vis.mod03.time.data, aes(x = t_since_in, y = visregFit), size = 1, col = 2)+
  geom_path(data = splits %>% filter(event %in% c("Start", "Split")), aes(x = t_since_in, y = fsize, group = fam), arrow = arrow(length = unit(0.02, "npc"), type = "closed"), lty = 1, lwd = 0.2, col = "grey30")+
  geom_text(data = splits %>% filter(fam %in% c("Ev","Wo","Ha","Na","Ti","Ma","Ro","An","Jo"), event == "Start"), aes(x=t_since_in, y = fsize, label = substr(fam,1,1), angle = 0), hjust = 1, vjust = 0, size = 3)+
   ylim(0,7.5)+ xlim(0,200)+
   geom_text(aes(x =180, y = 7, label = "{chi}^2==74.166"), parse = T, hjust = "inwards")+
  labs(list(x = "Days since arrival", y = "Family size", title = "a"))+
  theme_bw()+a

#dev.off()
```

# Family size ~ time: geese.org

```{r pred_famsize-time_gorg}
vis.famsize.time.gorg = visreg(mod03.gorg, "t_since_in", scale = "response",plot = F)

vis.famsize.time.gorg.data = vis.famsize.time.gorg$fit
```

```{r}
#postscript(file = "plot04famsize_time_gorg.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

p6 = ggplot()+
  geom_ribbon(data = vis.famsize.time.gorg.data, aes(x = t_since_in, ymin = visregLwr, ymax = visregUpr), fill = "grey90")+
  geom_line(data = vis.famsize.time.gorg.data, aes(x = t_since_in, y = visregFit), size = 1, col = 2)+
  labs(list(x = "Days since arrival", y = NULL, title = "b"))+
  geom_path(data = splits %>% filter(event %in% c("Start", "Split")), aes(x = t_since_in, y = fsize, group = fam), arrow = arrow(length = unit(0.02, "npc"), type = "closed"), lty = 1,lwd = 0.2, col = "grey30")+
  geom_text(data = splits %>% filter(fam %in% c("Ev","Wo","Ha","Na","Ti","Ma","Ro","An","Jo"), event == "Start"), aes(x=t_since_in, y = fsize, label = substr(fam,1,1), angle = 0), hjust = 1, vjust = 0, size = 3)+
   ylim(0,7.5)+ xlim(0,200)+
   geom_text(aes(x =180, y = 7, label = "{chi}^2==19.936"), parse = T, hjust = "inwards")+
  theme_bw()+b

#dev.off()
```

# Family size ~ flock size

```{r modpred_fam_flocksize}
#'predictions of family size from flocksizes
vis.mod03.famflock = visreg(mod03, "flocksize", scale = "response", plot = F)

#'make data frame
vis.mod03.famflock.data = vis.mod03.famflock$fit
```

```{r plot_mod}
#postscript(file = "plot08fam_flock.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

p7 = ggplot()+
  geom_point(data = fams.expand[runif(10000, 1, 34e3),], aes(x = (flocksize), y = famsize), size = 0.1)+
  geom_ribbon(data = vis.mod03.famflock.data, aes(x = flocksize, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = vis.mod03.famflock.data, aes(x = flocksize, y = visregFit), size = 1, col = 2)+
  labs(list(x = "Flock size", y = "Family size", title = "c"))+
  geom_text(data = NULL, aes(x = 1e4, y = 9.5, label = "{chi}^2==0.270"), hjust = "inward", parse = T)+
  ylim(0,10)+
  theme_bw()+a

#dev.off()
```

# Family size ~ p.index: Both datasets

```{r pred_famsize_pindex}
vis.famsize.pred = visreg(mod03, "p.index", scale = "response", plot = F)

vis.famsize.pred.data = vis.famsize.pred$fit

vis.famsize.gorg.pred = visreg(mod03.gorg, "p.index", scale = "response", plot = F)

vis.famsize.gorg.pred.data = vis.famsize.gorg.pred$fit

vis.famsize.gorg.s.pred = visreg(mod03.gorg.s, "p.index", scale = "response", plot = F)

vis.famsize.gorg.s.pred.data = vis.famsize.gorg.s.pred$fit
```

```{r plot_famsize_pred}
p8 = ggplot()+
  geom_point(data = geeseorg, aes(x = p.index, y = famsize), size = 0.1)+
  geom_ribbon(data = vis.famsize.pred.data, aes(x = p.index, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_ribbon(data = vis.famsize.gorg.pred.data, aes(x = p.index, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_ribbon(data = vis.famsize.gorg.s.pred.data, aes(x = p.index, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  #'dashed line, families in flocks
  geom_line(data = vis.famsize.pred.data, aes(x = p.index, y = visregFit), lty = 2, col = 2)+
  
  geom_line(data = vis.famsize.gorg.pred.data, aes(x = p.index, y = visregFit), lty = 1, col = 2)+
  geom_line(data = vis.famsize.gorg.s.pred.data, aes(x = p.index, y = visregFit), lty = 3, col = 2)+
  ylim(0,10)+
  #geom_text(aes(x = 1.25, y = c(2.5, 0.75), label = c("w/o unsuccessful pairs", "w/ unsuccessful pairs")))+
  geom_text(aes(x = 1.8, y = c(2.3, 1.3,0.75), label = c("1","2","3")))+
  geom_text(aes(x = 1.8, y = c(9.5,8.5,7.5), label = c("{chi[1]}^2==0.337", "{chi[2]}^2==0.1321","{chi[3]}^2==12.935")), hjust = "inward", parse = T)+
  labs(list(x = "Predation index", y = NULL, title = "d"))+
  theme_bw()+b
```


# Flocksize ~ distkol

```{r model_pred_flockmod}
#'get model pred
vis.flockmod = visreg(flockmod, "distkol", scale = "response", plot = F)

vis.flockmod.data = vis.flockmod$fit
```

```{r export_flock_lon_plot}
#postscript(file = "plot05flocksize_lon.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

p9 = ggplot()+
  geom_jitter(data = geese, aes(x = distkol, y = flocksize), size = 0.1, width = 5)+
  geom_ribbon(data = vis.flockmod.data, aes(x = distkol, ymin = visregLwr+10, ymax = visregUpr+10), fill = "grey90", alpha = 0.9)+
  geom_line(data = vis.flockmod.data, aes(x = distkol, y = visregFit), size = 1, col = 2)+
  ylim(0,5000)+
  labs(list(x = "Dist. to Kolguyev (km)", y = "Flock size", title = "a"))+
  geom_text(aes(x =3100, y = 4500, label = "{chi}^2==66599"), hjust = "inward", parse = T)+
  theme_bw()+a

#dev.off()
```

# Juvenile % ~ flocksize

```{r pred_gamjuvs_lon}
vis.juvs.flock = visreg(modjuvs, "flocksize", scale = "response",plot = F)

vis.juvs.flock.data = vis.juvs.flock$fit
```

```{r}
#postscript(file = "plot10juvprop_flocksize.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)
p10 = ggplot()+
  geom_point(data=geese, aes(x = flocksize, y = propjuv), size = 0.1)+
 geom_ribbon(data = vis.juvs.flock.data, aes(x = flocksize, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = vis.juvs.flock.data, aes(x = flocksize, y = visregFit), size = 1, col = 2)+
  ylim(0.05, 0.8)+
  xlim(0,5000)+
  labs(list(x = "Flock size", y = "Juvenile proportion", title = "b"))+
  geom_text(aes(x = 5000, y = 0.7, label = "{chi}^2==5.921"), hjust = "inward", parse = T)+
  theme_bw()+a

#dev.off()
```

# Juvenile % ~ winter

```{r pred_gamjuvs_time}
vis.juvs.time = visreg(modjuvs, "t_since_in", scale = "response",plot = F)

vis.juvs.time.data = vis.juvs.time$fit
```

```{r}
#postscript(file = "plot09juvprop_time.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

p11 = ggplot()+
  geom_point(data = geese, aes(x = t_since_in, y = propjuv), size = 0.1)+
  geom_ribbon(data = vis.juvs.time.data, aes(x = t_since_in, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = vis.juvs.time.data, aes(x = t_since_in, y = visregFit), size = 1, col = 2)+
  labs(list(x = "Days since arrival", y = "Juvenile proportion", title = "c"))+
  ylim(0.05, 0.8)+
  geom_text(aes(x = 200, y = 0.7, label = "{chi}^2==19.43"), hjust = "inward", parse = T)+
  theme_bw()+a

#dev.off()
```

# Plot : Juvenile % ~ p.index

```{r}
vis.juvprop.pred = visreg(modjuvs, "p.index", scale = "response", plot = F)

vis.juvprop.pred.data = vis.juvprop.pred$fit
```

```{r}
#postscript(file = "plot13juvprop_pred.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

p12 = ggplot()+
  geom_point(data = geese, aes(x=p.index, y = propjuv), size = 0.1)+
  geom_ribbon(data = vis.juvprop.pred.data, aes(x = p.index, ymin = visregLwr, ymax = visregUpr), fill= "grey90", alpha = 0.9)+
  geom_line(data = vis.juvprop.pred.data, aes(x = p.index, y = visregFit), size = 1, col=2)+
  labs(list(x = "Predation index", y = "Juvenile proportion", title = "d"))+
  geom_text(aes(x = 1.9, y = 0.7, label = "{chi}^2==0.021"), hjust = "inward", parse = T)+ylim(0,0.8)+
  theme_bw()+a

#dev.off()
```

# Predation plots

# Flock size ~ p.index

```{r pred_flocksize_pindex}
vis.flocksize.pred = visreg(flockmod, "p.index", scale = "response", plot = F)

vis.flocksize.pred.data = vis.flocksize.pred$fit
```

```{r}
#postscript(file = "plot13flocksize_pred.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

p11 = ggplot()+
  geom_point(data = geese, aes(x = p.index, y = flocksize), size = 0.1)+
  geom_ribbon(data = vis.flocksize.pred.data, aes(x = p.index, ymin = visregLwr, ymax = visregUpr), fill= "grey90", alpha = 0.9)+
  geom_line(data = vis.flocksize.pred.data, aes(x = p.index, y = visregFit), size = 1, col = 2)+
  labs(list(x = NULL, y = "Flock size", title = "a"))+
  ylim(0,9e3)+
  geom_text(aes(x = 1.5, y = 8500, label = "{Omega[0]}^2==0.99"), hjust = "inward", parse = T)+
  theme_bw()+a

#dev.off()
```

# Plot 13: Number of families ~ p.index: SOVON data

```{r}
vis.nfams.pred = visreg(famsmod, "p.index", scale = "response", plot = F)

vis.nfams.pred.data = vis.nfams.pred$fit
```

```{r}
p13 = ggplot()+
  geom_point(data = geese, aes(x = p.index, y = fams), size = 0.1)+
  geom_ribbon(data = vis.nfams.pred.data, aes(x = p.index, ymin = visregLwr, ymax = visregUpr), fill= "grey90", alpha = 0.9)+
  geom_line(data = vis.nfams.pred.data, aes(x = p.index, y = visregFit), size = 1, col = 2)+
  labs(list(x = "Predation index", y = "Successful families in flock", title = "c"))+
  ylim(0,190)+
  geom_text(aes(x = 1.5, y = 185, label = "{Omega[0]}^2==0.99"), hjust = "inward", parse = T)+
  theme_bw()+a
```

# Export plots
```{r}
save(p1,p2,p3,p4,p5,p6,p7,p8,p9, p10,p11, p12, file = "goose_dataplots.Rdata")
```

#Grid
```{r yr_month_heatmap, fig.fullwidth=TRUE, fig.height=6, fig.width=8, fig.cap= "Heatmap of samples in each month in each year between 2001 - 2011. Colour scale represents the number of records.", warning=FALSE, eval = TRUE}

library(gridExtra);library(ggplot2)

#'distance ~ fam size
pdf(file = "~/git/thesis/texts/dist_fsize.pdf")
grid.arrange(p1, p2, p3, p4)
dev.off()

#'family trends
pdf(file = "~/git/thesis/texts/fam_trends.pdf", height = 6, width = 8)
grid.arrange(p5, p6, p7, p8, ncol = 2)
dev.off()

#'flock trends
pdf(file = "~/git/thesis/texts/flock_trends.pdf", height = 6, width = 8)
grid.arrange(p9, p10, p11, p12, ncol = 2)
dev.off()
```

# Other plots

# Flocksize ~ time

* Significant effect, flocks are larger in the late winter

```{r model_pred_flockmod_time}
#'get model pred
vis.flockmod.days = visreg(flockmod, "t_since_in", scale = "response", plot = F)

vis.flockmod.days.data = vis.flockmod.days$fit

p6.res = vis.flockmod.days$res
```

```{r export_flock_time_plot}
#postscript(file = "plot06flocksize_days.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

#p8 = ggplot()+
  geom_point(data = geese, aes(x = t_since_in, y = flocksize), size = 0.1)+
  geom_ribbon(data = vis.flockmod.days.data, aes(x = t_since_in, ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = vis.flockmod.days.data, aes(x = t_since_in, y = visregFit), size = 1, col = 2)+
  labs(list(x = "Days since first arrivals", y = NULL, title = "b"))+
  ylim(0,5000)+
  theme_bw()+b

#dev.off()
```

# Number of families ~ flock size

```{r pred_famsmod}
visfamsmod = visreg(famsmod, "flocksize", plot = F, scale = "response")

visfamsmod.data = visfamsmod$fit
```

```{r plot_pred_famsmod}
#postscript(file = "plot07fams_in_flocks.eps", colormodel = "grey", horizontal = FALSE, onefile = FALSE, height = 4, width = 5)

#p9= ggplot()+
  geom_point(data = fams, aes(x = log(flocksize), y = fams), size = 0.1)+
  geom_ribbon(data = visfamsmod.data, aes(x = log(flocksize), ymin = visregLwr, ymax = visregUpr), fill = "grey90", alpha = 0.9)+
  geom_line(data = visfamsmod.data, aes(x = log(flocksize), y = visregFit), size = 1, col = 2)+ ylim(0,2e2)+
  labs(list(x = "Flock size", y = "Number of families", title = "c"))+
  geom_text(aes(x =1.5, y = 190, label = "{chi}^2==7769"), hjust = "inward", parse = T)+
  theme_bw()+a

#dev.off()
```
